---
title: "Report"
execute: 
  echo: false      # change to true to show the code
code-fold: true  # change to true to fold the code chunks
---

Your report will go here.

```{r}
# this is a test code
```

How Pumpkin Growing Power Differs Across the World.

Every fall, growers all over the world compete to grow the heaviest pumpkins. But not all pumpkins, or all regions, are created equal. Our analysis finds that pumpkin weight varies significantly based on genetics, location, and even yearly conditions.

We analyzed competition data from the Great Pumpkin Commonwealth (GPC) from 2013 to 2020. We both focus on the pumpkin growth the global scale and specifically about the the United States, where the majority of entries are recorded. Using over 6,000 entries, we explored how seed genetics and regional growing conditions impact pumpkin weight.

Genetic Factors: Seed Mothers and Pollinator Fathers

Genetics plays a major role in how heavy a pumpkin can grow. We first looked at the "seed mother" lineage—the maternal side of the pumpkin family.

```{r}
tuesdata <- tidytuesdayR::tt_load('2021-10-19')
tuesdata <- tidytuesdayR::tt_load(2021, week = 43)

pumpkins <- tuesdata$pumpkins

# Or read in the data manually

pumpkins <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2021/2021-10-19/pumpkins.csv')
library(dplyr)
library(stringr)
library(ggplot2)

pumpkin_clean <- pumpkins %>%
  mutate(
    weight_lbs = as.numeric(weight_lbs),
    seed_mother = as.factor(seed_mother),
    pollinator_father = as.factor(pollinator_father)
  ) %>%
  filter(!is.na(weight_lbs), !is.na(seed_mother), !is.na(pollinator_father))
pumpkin_clean %>%
  filter(seed_mother %in% names(sort(table(seed_mother), decreasing = TRUE)[1:10])) %>%
  ggplot(aes(x = seed_mother, y = weight_lbs)) +
  geom_boxplot(fill = "#FFA500", color = "#8B4513") +
  labs(title = "Pumpkin Weight by Top 10 Seed Mothers", x = "Seed Mother", y = "Weight (lbs)",caption = "Data source: GPC Competition Data (2013-2020)") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```


The data shows that pumpkins from seed mothers like "2008 Neptune" and "2145 McMullen" consistently produce heavier entries, while others like "81 Wolf" have lower median weights. This highlights that superior seed lines can offer growers a significant advantage.

The pollinator father's side tells a similar story.

```{r}
library(ggplot2)

pumpkin_clean %>%
  filter(pollinator_father %in% names(sort(table(pollinator_father), decreasing = TRUE)[1:10])) %>%
  ggplot(aes(x = pollinator_father, y = weight_lbs)) +
  geom_boxplot(fill = "#FFA500", color = "#8B4513") +
  labs(title = "Pumpkin Weight by Top 10 pollinator_father", x = "pollinator_father", y = "Weight (lbs)", caption = "Data source: GPC Competition Data (2013-2020)") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```


Notably, pumpkins pollinated by "2009 Wallace" lines stood out with much heavier median weights, compared to more generic "self" or "open" pollination sources. The importance of strategic breeding is clear: winning pumpkins often come from carefully selected parentage.

Regional Trends: How Pumpkin Weight Changes Over Time

Location matters too. Environmental factors such as soil quality, temperature, and growing culture differ across regions. We zoomed into the six U.S. states and provinces with the most competition entries: California, Michigan, Ohio, Ontario, Wisconsin, and a general "Other" category.

```{r}
library(tidyverse)

pumpkins <- pumpkins %>%
  mutate(year = as.numeric(str_extract(id, "^\\d{4}"))) %>%
  filter(!is.na(state_prov), !is.na(weight_lbs))


top_states <- pumpkins %>%
  count(state_prov, sort = TRUE) %>%
  slice_head(n = 6) %>%
  pull(state_prov)


pumpkins_avg <- pumpkins %>%
  filter(state_prov %in% top_states) %>%
  group_by(state_prov, year) %>%
  summarise(avg_weight = mean(as.numeric(weight_lbs), na.rm = TRUE)) %>%
  ungroup()


ggplot(pumpkins_avg, aes(x = year, y = avg_weight)) +
  geom_line(color = "#FFA500", size = 1.2) +  # 
  geom_point(color = "black", size = 2) +      # 
  facet_wrap(~ state_prov) +
  labs(
    title = "Average Pumpkin Weight Over Time by State (Top 6 States)",
    x = "Year",
    y = "Average Weight (lbs)",
    caption = "Data source: GPC Competition Data (2013-2020)"
  ) +
  theme_minimal() +
  theme(
    strip.text = element_text(size = 10),
    plot.title = element_text(hjust = 0.5, face = "bold"),
    plot.caption = element_text(hjust = 0.5, size = 8)
  )

```


From 2013 to 2020, no state showed dramatic year-over-year growth, but there were fluctuations. California showed relatively higher weight spikes around 2017, while Ontario and Wisconsin had steadier, flatter trends. Michigan and Ohio displayed more variability, hinting at changing growing conditions or participant turnover.

Geographic Distribution: Where the Heaviest Pumpkins Are Grown

To visualize the regional differences more clearly, we mapped the average pumpkin weight across U.S. states.
```{r}
library(tidyverse)
library(maps)
library(ggthemes)


pumpkins_us <- pumpkins %>%
  filter(country == "United States", !is.na(state_prov), !is.na(weight_lbs)) %>%
  mutate(state_prov = tolower(state_prov))


state_avg_weight <- pumpkins_us %>%
  group_by(state_prov) %>%
  summarise(avg_weight = mean(as.numeric(weight_lbs), na.rm = TRUE)) %>%
  ungroup()


us_map <- map_data("state")

us_map_data <- us_map %>%
  left_join(state_avg_weight, by = c("region" = "state_prov"))


ggplot(us_map_data, aes(x = long, y = lat, group = group, fill = avg_weight)) +
  geom_polygon(color = "white") +
  coord_fixed(1.3) +
  scale_fill_gradient(
    low = "#FFF5E1",   # 更接近白色的浅米色
    high = "#FF4500",  # 深橙红色（接近南瓜皮深色）
    na.value = "grey90",
    name = "Avg Weight (lbs)"
  ) +
  labs(
    title = "Average Pumpkin Weight by U.S. State",
    subtitle = "Based on GPC competition data (2013-2020)",
    caption = "Data source: TidyTuesday / GPC"
  ) +
  theme_minimal() +
  theme(
    axis.title = element_blank(),
    axis.text = element_blank(),
    axis.ticks = element_blank(),
    panel.grid = element_blank(),
    plot.title = element_text(size = 14, hjust = 0.5, face = "bold"),
    plot.subtitle = element_text(size = 10, hjust = 0.5),
    plot.caption = element_text(hjust = 0.5, size = 8),
    legend.position = "bottom",
    panel.background = element_rect(fill = "white", color = NA),
    plot.background = element_rect(fill = "white", color = NA)
  )

```



The Midwest, especially states like Ohio and Michigan, consistently produced heavier pumpkins compared to southern or coastal states. This suggests that cooler climates, specific soil types, or stronger pumpkin-growing traditions could favor giant pumpkin cultivation.

Conclusion: Winning Takes Both Genes and Geography

Our analysis shows that growing a giant pumpkin isn't just a matter of luck. Genetics—both seed mothers and pollinator fathers—have a major influence on potential size. Meanwhile, geography—the local environment and growing conditions—also plays a critical role.

In short, the best pumpkin growers combine the right seeds with the right place. As competition intensifies, understanding and optimizing both factors will be key for those hoping to set future records.

Data source: TidyTuesday / GPC

